Complete Real-Time Liver Model including Glisson Capsule, Vascular Network and Parenchyma

### 2013-02-28 Doc/Miccai2013/miccai2013.tex
ADD: introduction
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@526 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

%%% Comments
\usepackage{color}
\newcommand{\TG}[1]{{\color{blue}\textbf{TG: #1}}}
\newcommand{\CD}[1]{{\color{green}\textbf{CD: #1}}}
%%%

@@

\section{Introduction} %{{{

Medical simulation is expected to play an integral part in many aspects of
medical practice. From diagnosis, through planning or training to computer
assisted intervention. A good medical model has to be simple yet efficient
to be really usable.
\CD{develop the applications for which it has to be real-time...}

\CD{First... explain the biomechanics of the liver (for instance: 
\textit{
The liver is a complex organ composed by three main structures: the parenchyma (soft tissue, that have a hyper-visco-elastic behavior) is coupled with a dense vascular network (tree-like tubular structure with a stiffer behavior) and is surrounded by the Glisson capsule  (a stiff collagenous membrane).
It illustrates the need of relying on both volume, surface and line finite elements and the need of proposing coupling methods between them.}
 }

\CD{Yet, many work on liver modeling only consider the parenchyma...}
The liver is usually modeled as a homogeneous isotropic (visco-)elastic
material \TG{refs}.% It is however non-homogeneous and modeling it as
%homogeneous necessarily requires complex non-linear material descriptions.
Recently
Peterl\'{i}k et al. \cite{Peterlik2012} showed the importance of the vascular
structures on the liver deformation.

Another often overlooked component is the Glisson's capsule.
Biomechanical measurements report that the capsule has very small
thickness (approx. 20 $\mu$m for porcine liver \cite{Umale2011}).
On the other hand it has in three orders of magnitude higher stiffness than
the parenchyma.
Because of this it plays an important role in local
deformations of the liver \cite{Ahn2010,Hollenstein2006}.
It was reported that evaluation of the liver as a homogeneous material
without capsule can lead to overestimation of local deformations up to the factor of 3
\cite{Hollenstein2006}.

While the influence on local deformation was reported, no experimental
results show whether the capsule also plays an important role on the global
deformations of the liver. We propose a composite model that models the
capsule with membrane FEM elements. We show that the model is able to
simulate expected local behaviour reported in literature. Further more we
show that the capsule plays also an important role in the deformation on
global scale.

\TG{summary of following sections.}

>>>>

\def\wrt{w.\,r.\,t.}
\def\eg{e.\,g.}
\def\ie{i.\,e.}
\def\Dash{\nobreak\,---\penalty-500\,}

\def\TODO{{\color{red}{TODO}}}

%%% Comments
\usepackage{color}
\newcommand{\TG}[1]{{\color{blue}\textbf{TG: #1}}}
\newcommand{\CD}[1]{{\color{green}\textbf{CD: #1}}}
\newcommand{\IP}[1]{{\color{cyan}\textbf{IP: #1}}}
%%%

@@

\section{Introduction} 
\IP{I think we should provide some introductory motivation and broader context of the reasarch. I tried this:}

In 2009, nearly 100,000 European citizens died of cirrhosis or liver cancer. 
According to the statistics, nearly 100,000 European citizens die of cirrhosis or liver cancer each year. 
Although new methods such as radio-frequency ablation or cryo-ablation used in interventional radiology 
represent promising improvements, surgery remains the option that offers the foremost success rate against these pathologies. 
Nevertheless, surgery is not always performed due to several limitations, in particular the determination 
of accurate eligibility criteria for the patient. 
In this context, pre-operational planning becomes a crucial task having a significant impact on the treatment.

Computer-aided physics-based medical simulation has proven to be an extremely useful technique in the area of medical training. 
Nevertheless, whereas generic models are usually required in training simulators, accurate patient-specific modelling
becomes necessary as soon as computer simulation is to be employed in the pre-operative planning. At the same time, 
interactivity of such models  remains an important aspect, requiring real-time simulation which is often difficult to 
achieve given the complexity of soft tissues.

In case of human liver, the task of real-time accurate modeling is challenging, mainly because of the complex structure 
of this organ composed of three main constituents: \emph{parenchyma}, \emph{vascular networks} and \emph{Glisson's capsule}.
The parenchyma is certainly the most studied component of the liver; today the researchers agree on hyperelastic 
properties of the tissue, for which the mechanical parameters have been reported for example in~\cite{}. 
Several methods have been proposed to model the hyperelastic behaviour at real-time rates, such multiplicative Jacobian decomposition
introduced in~\cite{stephanie2008}.

The mechanical importance of the vascular structures in liver is studied in~\cite{peterlik2012}. It shows that that the 
influence of vessels on the mechanical behaviour of the liver is significant, mainly if large deformations occur. 
Since detailed modeling of the vessels would be extremely costly (mainly because of the thickness of the vessel wall 
being less than 250\,{$\mu$}m), the authors propose a composite model allowing for real-time simulation of entire liver
with two vascular trees corresponding to portal and hepatic veins.

The final constituent represented by Glisson's capsule have been already studied in the literature. 
Extensive experiments on porcine liver has been published in~\cite{Sagar}. The measurements show that although being very 
thin (10--20$\mu m$), the Young's modulus of the capsule (reported as $8.22\pm3.42$\,MPa) exceeds the values for the parenchyma by three orders of magnitude,
suggesting the mechanical influence of the membrane on the liver behaviour.

In~\cite{Hollenstein}, a local influence of the capsule has been measured using a special aspiration device. The study was then repeated 
in vivo on human patients during the operation, confirming the mechanical importance of the membrane~\cite{Ahn,Nava}.
To our best knowledge, on attempt has been made to demonstrate the role of the Glisson's capsule on the global behaviour of 
the liver, i.e. during large deformations of the organ.

The main contribution of the paper is twofold: first, we present an efficient real-time model of the liver, taking into account 
the parenchyma, vascularization and Glisson's capsule. The model is based on three different finite element representations for each constituent,
linked together via mechanical coupling. We show that the model mimics the local experiments descibed in~\cite{Hollenstein}.

Second, we use the complete model of the liver to demonstrate the influence of the Glisson's
capsule via simulation: using a specific model of a pig liver built from CT contrast-enhanced data, we show that there is a significant 
difference in the response of the model with and without the capsule in case when the liver undergoes large deformations.

The paper is organized as follows: \TODO.

%%Medical simulation is expected to play an integral part in many aspects of
%%medical practice. From diagnosis, through planning or training to computer
%%assisted intervention. A good medical model has to be simple yet efficient
%%to be really usable.
%\CD{develop the applications for which it has to be real-time...}
%
%\CD{First... explain the biomechanics of the liver (for instance: 
%\textit{
%The liver is a complex organ composed by three main structures: the parenchyma (soft tissue, that have a hyper-visco-elastic behavior) is coupled with a dense vascular network (tree-like tubular structure with a stiffer behavior) and is surrounded by the Glisson capsule  (a stiff collagenous membrane).
%It illustrates the need of relying on both volume, surface and line finite elements and the need of proposing coupling methods between them.}
% }
%
%\CD{Yet, many work on liver modeling only consider the parenchyma...}
%The liver is usually modeled as a homogeneous isotropic (visco-)elastic
%material \TG{refs}.% It is however non-homogeneous and modeling it as
%%homogeneous necessarily requires complex non-linear material descriptions.
%Recently
%Peterl\'{i}k et al. \cite{Peterlik2012} showed the importance of the vascular
%structures on the liver deformation.
%
%Another often overlooked component is the Glisson's capsule.
%Biomechanical measurements report that the capsule has very small
%thickness (approx. 20 $\mu$m for porcine liver \cite{Umale2011}).
%On the other hand it has in three orders of magnitude higher stiffness than
%the parenchyma.
%Because of this it plays an important role in local
%deformations of the liver \cite{Ahn2010,Hollenstein2006}.
%It was reported that evaluation of the liver as a homogeneous material
%without capsule can lead to overestimation of local deformations up to the factor of 3
%\cite{Hollenstein2006}.
%
%While the influence on local deformation was reported, no experimental
%results show whether the capsule also plays an important role on the global
%deformations of the liver. We propose a composite model that models the
%capsule with membrane FEM elements. We show that the model is able to
%simulate expected local behaviour reported in literature. Further more we
%show that the capsule plays also an important role in the deformation on
%global scale.
%
\TG{summary of following sections.}
